---
title: "Titanic (Kaggle Competition): Feature Engineering"
output: 
  # Output as a previewable HTML Notebook.
  html_notebook:
    theme: united
    highlight: tango
    df_print: paged
    toc: true
    toc_depth: 3
  # Output as an rmarkdown::pdf_document()
  pdf_document:
    keep_tex: true
    highlight: tango
    fig_caption: true
    df_print: kable
    toc: true
    toc_depth: 3
# pandoc -> pdfLaTeX rendering options:
fontsize: 11pt
geometry:
  - margin=1in
  - heightrounded
documentclass: scrartcl
papersize: a4
urlcolor: violet
toccolor: blue
---

## Feature Engineering
### `Deck` (extracted from `Cabin`)

### `TicketPrefix` (extracted from `Ticket`)

### `TicketNumber` (extracted from `Ticket`)


### `Fare_Rounded` (extracted from `Fare`)

#### Summary

```{r feature-fare-range}
Titanic.df %<>%
    dplyr::mutate(Fare_Range = cut(Fare, 
                                    breaks = c(-Inf, seq(0, 550, by = 50), Inf),
                                    right = TRUE,
                                    ordered_result = TRUE),
                  Fare_Rounded = round(Fare / 100, 1) * 100)

Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-fare-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, Fare_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))

Titanic.df %>%
    dplyr::group_by(Subset, Fare_Rounded) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

```{r feature-summary-fare-range}
Titanic.df %>%
  sjstats::means_by_group("Fare_Range", "Survived")

Titanic.df %>%
  sjstats::means_by_group("Fare_Rounded", "Survived")
```

### `InfChiAdu` (extracted from `Age`)

Factor representing if the passenger is an infant, child, or adult. Useful for interpretation and helpful for additional feature engineering.

### `WithParent` (extracted from `InfChiAdu` and `Parch`)

If `InfChiAd` is either 1 or 2 and if Parch is 0.... child is with nanny.

### `WithChild` (extracted from `InfChiAdu` and `Ticket`)

### `SibSp_Range` (extracted from `SibSp`)

Our previous review of `SibSp` suggested that more than 90% of the observed sample has 0 siblings and/or spouses on board at the time of the shipwreck. Can we filter this down into a smaller amount of factors that better describe outcomes?

#### Summary

```{r feature-sibsp-range}
Titanic.df %<>%
    dplyr::mutate(SibSp_Range = cut(SibSp, 
                                    breaks = c(-Inf, 0, 1, Inf),
                                    labels = c("0", 
                                               "1", 
                                               "2+"),
                                    right = TRUE,
                                    ordered_result = TRUE))
Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-sibsp-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, SibSp_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

#### By Response

```{r eda-kruskal-sibsp-range-response}
(sjstats::means_by_group(Titanic.df, "Survived", "SibSp_Range"))
(kt <- kruskal.test(Survived ~ SibSp_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-sibsp-range-response}
TSRS <- xtabs(~ Survived + SibSp_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TSRS)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TSRS), 2))
```

```{r eda-mosaic-sibsp-range-response}
vcd::mosaic(TSRS, shade=TRUE, legend=TRUE)
vcd::assoc(TSRS, shade=TRUE)
```

#### By `Sex`

```{r eda-kruskal-sibsp-range-gender}
(sjstats::means_by_group(Titanic.df, "SibSp_Range", "Sex"))
(kt <- kruskal.test(Sex ~ SibSp_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-sibsp-range-gender}
TSRG <- xtabs(~ Sex + SibSp_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TSRG)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TSRG), 2))
```

```{r eda-mosaic-sibsp-range-gender}
vcd::mosaic(TSRG, shade=TRUE, legend=TRUE)
vcd::assoc(TSRG, shade=TRUE)
```

#### By `Age`

```{r eda-kruskal-sibsp-range-age}
(sjstats::means_by_group(Titanic.df, "Age", "SibSp_Range"))
(r <- cor.test(~ Age + as.numeric(SibSp_Range), data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", r$p.value < 0.05)))
```

```{r eda-boxplot-sibsp-range-age, fig.width=6}
Titanic.df %>%
  dplyr::select(Subset, Survived, SibSp_Range, Age) %>%
  ggplot2::ggplot(ggplot2::aes(x = SibSp_Range, y = Age, fill = SibSp_Range, group = SibSp_Range)) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(title = "Boxplot of Age by Sibling/Spouse Count", x = "Number of Siblings and/or Spouses", y = "Passenger Age")
```


### `Parch_Range` (extracted from `Parch`)

#### Summary

```{r feature-parch-range}
Titanic.df %<>%
    dplyr::mutate(Parch_Range = cut(Parch, 
                                    breaks = c(-Inf, 0, 1, Inf),
                                    labels = c("0", 
                                               "1", 
                                               "2+"),
                                    right = TRUE,
                                    ordered_result = TRUE))
Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-parch-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, Parch_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

#### By Response

```{r eda-kruskal-parch-range-response}
(sjstats::means_by_group(Titanic.df, "Survived", "Parch_Range"))
(kt <- kruskal.test(Survived ~ Parch_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-parch-range-response}
TPCS <- xtabs(~ Survived + Parch_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TPCS)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TPCS), 2))
```

```{r eda-mosaic-parch-range-response}
vcd::mosaic(TPCS, shade=TRUE, legend=TRUE)
vcd::assoc(TPCS, shade=TRUE)
```

#### By `Sex`

```{r eda-kruskal-parch-range-gender}
(sjstats::means_by_group(Titanic.df, "Parch_Range", "Sex"))
(kt <- kruskal.test(Sex ~ Parch_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-parch-range-gender}
TPCG <- xtabs(~ Sex + Parch_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TPCG)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TPCG), 2))
```

```{r eda-mosaic-parch-range-gender}
vcd::mosaic(TPCG, shade=TRUE, legend=TRUE)
vcd::assoc(TPCG, shade=TRUE)
```

#### By `Age`

```{r eda-kruskal-parch-range-age}
(sjstats::means_by_group(Titanic.df, "Age", "Parch_Range"))
(r <- cor.test(~ Age + as.numeric(Parch_Range), data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", r$p.value < 0.05)))
```

```{r eda-boxplot-parch-range-age, fig.width=6}
Titanic.df %>%
  dplyr::select(Subset, Survived, Parch_Range, Age) %>%
  ggplot2::ggplot(ggplot2::aes(x = Parch_Range, y = Age, fill = Parch_Range, group = Parch_Range)) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(title = "Boxplot of Age by Parent/Child Count", x = "Number of Parents and/or Children", y = "Passenger Age")
```


### Latent Features in `Name`

```{r feature-name}
# Prepare data frame for extraction from the `Name` feature.
(Titanic.name <- dplyr::select(Titanic.df, PassengerId, Name, Subset))
```

```{r feature-name-explore}
# PRCE format Regex for formatting the names.
Titanic.name.format <- "^(.+?(?=,))(?>,\\s+)(.+?\\.)(?>\\s+)(.*?(?=\\t|$|\\s\\())(?|(?=\\t|$)()(?:.*?(?=$))|(?:\\s\\(?)(.*?(?=\\)))(?:.*?(?=$)))"
Titanic.name.matches <- regexec(
    pattern = Titanic.name.format,
    text = Titanic.name$Name,
    perl = TRUE # For PRCE conditionals.
)

# Generate the matches.
Titanic.name.components <- c("Source", "Surname", "Title", "Given Name (B)", "Given Name (A)")
(Titanic.name.features <- Titanic.name %>%
    dplyr::pull(Name) %>%
    regmatches(Titanic.name.matches) %>%
    as.data.frame(row.names = Titanic.name.components, col.names = Titanic.name$PassengerId) %>%
    tibble::rownames_to_column(var = "Component") %>%
    tidyr::pivot_longer(cols = -Component, names_to = "PassengerId", values_to = "Value") %>%
    dplyr::arrange(PassengerId, desc(Component)) %>%
    dplyr::mutate(PassengerId = as.integer(sub("X", "", PassengerId))))
```


### `Title` (extracted from `Name`)

```{r feature-name-title-extract}
# Prepare data frame for extraction from the `Name` feature.



```

### `Martial Status` (extracted from `Name`)


### `Forename` (extracted from `Name`)

### `Middlename` (extracted from `Name`)

<!-- Additional names besides first name -->

### `Surname` (extracted from `Name`)

