---
title: "Titanic (Kaggle Competition): Feature Engineering"
output: 
  # Output as a previewable HTML Notebook.
  html_notebook:
    theme: united
    highlight: tango
    df_print: paged
    toc: true
    toc_depth: 3
  # Output as an rmarkdown::pdf_document()
  pdf_document:
    keep_tex: true
    highlight: tango
    fig_caption: true
    df_print: kable
    toc: true
    toc_depth: 3
# pandoc -> pdfLaTeX rendering options:
fontsize: 11pt
geometry:
  - margin=1in
  - heightrounded
documentclass: scrartcl
papersize: a4
urlcolor: violet
toccolor: blue
---

```{r setup, include=FALSE}
# Set the root directory.
knitr::opts_knit$set(root.dir = here::here(""))

# Set the chunk options.
knitr::opts_chunk$set(
    ## ---- Formatting Options ----
    comment = "",
    include = TRUE,
    collapse = TRUE,
    echo = TRUE,
    strip.white = TRUE,
    
    ## ---- Output Options ----
    message = TRUE,
    warning = TRUE,
    error = TRUE,
    results = 'markup',
    
    ## ---- Display Options ----
    fig.height = 4,
    fig.width = 6
)

# Source loader file.
source(here::here("R/loader.R"))
load.data()
```

This report contains feature engineering details about the Titanic prediction models.

## Setup

Our investigation makes use of several R packages that made it easier to summarize, visualize, manipulate, model, and impute missing data. These packages are specified below.

```{r load-packages, include=TRUE, echo=FALSE}
library(magrittr) # Attached for the `%>%` (pipe) and `extract` functions.

loadNamespace("renv")    # Project dependency tracker for collaboration.
loadNamespace("here")    # Used for obtaining paths relative to project root.
loadNamespace("readr")   # Tidyverse data reader.
loadNamespace("tidyr")   # Tidyverse data wrangler.
loadNamespace("qqplotr") # For plotting quantile-quantile plots.
loadNamespace("ggmosaic") # For plotting mosaic plots.
loadNamespace("visdat")  # For visualization of data.
loadNamespace("ltm")     # For EDA.
loadNamespace("sjstats") # For EDA.
loadNamespace("vcd")     # For EDA.
loadNamespace("skimr")   # For finding statistical summaries.
loadNamespace("naniar")  # Handle NA values in a tabular format.
loadNamespace("simputation")  # Missing data imputation.
loadNamespace("e1071")   # For classification.
loadNamespace("randomForest") # For classification.
loadNamespace("neuralnet")    # For deep learning.
```

```{r feature-utilities}
make.names_range <- function(prefix = "X", 
                             range = 1:5,
                             suffix = "", 
                             sep = "",
                             width = 1) {
    width.format <- sprintf("%%0%dd", width)
    name.format <- paste(c(prefix, width.format, suffix), collapse = sep)
    return(c(sprintf(name.format, range)))
}
make.names_range("Cabin", 1:4, "Deck", sep = ".", width = 1)
```


## Feature Engineering

### Latent Features in `Cabin`

```{r feature-cabin}
# Neat little extraction function for handling Cabin numbers.
(Titanic.cabins <- Titanic.df %>%
    dplyr::select(PassengerId, Subset, Cabin) %>%
    dplyr::mutate(Cabin.Known = !is.na(Cabin)) %>%
    dplyr::mutate(
      Cabin = stringr::str_split(Cabin, pattern = "\\s")
      ) %>%
    tidyr::unnest_wider(Cabin, names_sep = ".") %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      Cabin.Count = sum(!is.na(dplyr::c_across(dplyr::num_range("Cabin.", 1:4))))
      ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(dplyr::across(
        dplyr::num_range("Cabin.", 1:4), 
        list(
          Deck = ~ stringr::str_extract(.x, ".+?"),
          Number = ~ stringr::str_extract(.x, "\\d+")
        ), 
        .names = "{.col}.{.fn}"
      )) %>%
    dplyr::mutate(dplyr::across(
        dplyr::starts_with("Cabin.") & dplyr::ends_with(".Deck"),
        ~ factor(.x, levels = c("A", "B", "C", "D", "E", "F", "G"), exclude = NA)
      )) %>%
    dplyr::mutate(dplyr::across(
      dplyr::starts_with("Cabin.") & dplyr::ends_with(".Number"),
      ~ as.integer(.x)
    )))
```

```{r feature-cabin-summary}
# Overview.
Titanic.cabins %>% skimr::skim()
```

#### `Deck` (extracted from `Cabin`)

```{r feature-cabin-summary-combinations-deck-subset}
# Counts for unique Cabin combinations (by Subset)
Titanic.cabins %>%
  dplyr::select(Subset, 
                make.names_range("Cabin", 1:4, "Deck", sep =".")) %>%
  dplyr::group_by_all() %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.1.Deck, 
                 Cabin.2.Deck, 
                 Cabin.3.Deck, 
                 Cabin.4.Deck, desc(Subset)) %>% print() %>% visdat::vis_miss()
```


```{r feature-cabin-summary-combinations-deck}
# Counts for unique Cabin combinations
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Deck", sep =".")) %>%
  dplyr::group_by_all() %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.1.Deck, 
                 Cabin.2.Deck, 
                 Cabin.3.Deck, 
                 Cabin.4.Deck) %>% print() %>%
  visdat::vis_miss()
```

```{r feature-cabin-summary-deck-subset}
# Counts for Deck counts (by Subset)
Titanic.cabins %>%
  dplyr::select(Subset,
                make.names_range("Cabin", 1:4, "Deck", sep =".")) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>%
  dplyr::mutate(
    dplyr::across(Cabin.1.Deck, ~ tidyr::replace_na(.x, "?"))) %>%
  tidyr::pivot_longer(Cabin.1.Deck:Cabin.4.Deck, 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Deck", values_drop_na = TRUE) %>%
  dplyr::group_by(Subset, Cabin.Deck) %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.Deck, desc(Count)) %>% print() %>%
  dplyr::slice(-c(1:2)) %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Deck, y = Count, fill = Cabin.Deck)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::facet_grid(~Subset) +
  ggplot2::theme(legend.position = "none")
```

```{r feature-cabin-summary-deck}
# Counts for Deck counts
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Deck", sep =".")) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>%
  dplyr::mutate(
    dplyr::across(Cabin.1.Deck, ~ tidyr::replace_na(.x, "?"))) %>%
  tidyr::pivot_longer(Cabin.1.Deck:Cabin.4.Deck, 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Deck", values_drop_na = TRUE) %>%
  dplyr::group_by(Cabin.Deck) %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.Deck, desc(Count)) %>%
  dplyr::slice(-1) %>% print() %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Deck, y = Count, fill = Cabin.Deck)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::theme(legend.position = "none")
```

```{r feature-cabin-summary-deck-missing}
# Counts for Deck counts
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Deck", sep =".")) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) %>%
  dplyr::mutate(
    dplyr::across(Cabin.1.Deck, ~ tidyr::replace_na(.x, "?"))) %>%
  tidyr::pivot_longer(Cabin.1.Deck:Cabin.4.Deck, 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Deck", values_drop_na = TRUE) %>%
  dplyr::group_by(Cabin.Deck) %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.Deck, desc(Count)) %>% print() %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Deck, y = Count, fill = Cabin.Deck)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::theme(legend.position = "none")
```


#### `CabinNumber` (extracted from `Cabin`)

```{r feature-cabin-summary-combinations-number-subset}
# Counts for unique Cabin combinations (by Subset)
Titanic.cabins %>%
  dplyr::select(Subset, 
                make.names_range("Cabin", 1:4, "Number", sep =".")) %>%
  dplyr::group_by_all() %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.1.Number, 
                 Cabin.2.Number, 
                 Cabin.3.Number, 
                 Cabin.4.Number, desc(Subset)) %>% print() %>% visdat::vis_miss()
```


```{r feature-cabin-summary-combinations-number}
# Counts for unique Cabin combinations
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Number", sep =".")) %>%
  dplyr::group_by_all() %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(desc(Count),
                 Cabin.1.Number, 
                 Cabin.2.Number, 
                 Cabin.3.Number, 
                 Cabin.4.Number) %>% print() %>%
  visdat::vis_miss()
```

```{r feature-cabin-summary-number-subset}
# Counts for Number counts (by Subset)
Titanic.cabins %>%
  dplyr::select(Subset,
                make.names_range("Cabin", 1:4, "Number", sep =".")) %>%
  dplyr::mutate(dplyr::across(Cabin.1.Number:Cabin.4.Number, as.integer)) %>%
  dplyr::mutate(
    dplyr::across(Cabin.1.Number, ~ tidyr::replace_na(.x, -1))) %>%
  tidyr::pivot_longer(Cabin.1.Number:Cabin.4.Number, 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Number", values_drop_na = TRUE) %>%
  dplyr::group_by(Subset, Cabin.Number) %>%
  dplyr::summarise(Count = dplyr::n(), .groups = "drop") %>%
  dplyr::arrange(Cabin.Number, desc(Count)) %>% print() %>%
  dplyr::slice(-c(1:2)) %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Number, y = Count, fill = Subset)) +
  ggplot2::geom_histogram(stat = "identity")
```

```{r feature-cabin-summary-number-histogram}
# Counts for Number counts
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Number", sep =".")) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.integer)) %>%
  tidyr::pivot_longer(c(Cabin.1.Number:Cabin.4.Number), 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Number", values_drop_na = TRUE) %>%
  dplyr::arrange(Cabin.Set) %>% print() %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Set)) +
  ggplot2::geom_histogram(stat = "count")
```

```{r feature-cabin-summary-number-missing}
# Counts for Number counts
Titanic.cabins %>%
  dplyr::select(make.names_range("Cabin", 1:4, "Number", sep =".")) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), as.integer)) %>%
  dplyr::mutate(
    dplyr::across(Cabin.1.Number, ~ tidyr::replace_na(.x, -1))) %>%
  tidyr::pivot_longer(c(Cabin.1.Number:Cabin.4.Number), 
                      names_to = "Cabin.Set", 
                      values_to = "Cabin.Number", values_drop_na = TRUE) %>%
  dplyr::arrange(Cabin.Set) %>% print() %>%
  ggplot2::ggplot(ggplot2::aes(x = Cabin.Set)) +
  ggplot2::geom_histogram(stat = "count")
```




### `TicketPrefix` (extracted from `Ticket`)

### `TicketNumber` (extracted from `Ticket`)


### `Fare_Rounded` (extracted from `Fare`)

#### Summary

```{r feature-fare-range}
Titanic.df %<>%
    dplyr::mutate(Fare_Range = cut(Fare, 
                                    breaks = c(-Inf, seq(0, 550, by = 50), Inf),
                                    right = TRUE,
                                    ordered_result = TRUE),
                  Fare_Rounded = round(Fare / 100, 1) * 100)

Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-fare-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, Fare_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))

Titanic.df %>%
    dplyr::group_by(Subset, Fare_Rounded) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

```{r feature-summary-fare-range}
Titanic.df %>%
  sjstats::means_by_group("Fare_Range", "Survived")

Titanic.df %>%
  sjstats::means_by_group("Fare_Rounded", "Survived")
```

### `InfChiAdu` (extracted from `Age`)

Factor representing if the passenger is an infant, child, or adult. Useful for interpretation and helpful for additional feature engineering.

### `WithParent` (extracted from `InfChiAdu` and `Parch`)

If `InfChiAd` is either 1 or 2 and if Parch is 0.... child is with nanny.

### `WithChild` (extracted from `InfChiAdu` and `Ticket`)

### `SibSp_Range` (extracted from `SibSp`)

Our previous review of `SibSp` suggested that more than 90% of the observed sample has 0 siblings and/or spouses on board at the time of the shipwreck. Can we filter this down into a smaller amount of factors that better describe outcomes?

#### Summary

```{r feature-sibsp-range}
Titanic.df %<>%
    dplyr::mutate(SibSp_Range = cut(SibSp, 
                                    breaks = c(-Inf, 0, 1, Inf),
                                    labels = c("0", 
                                               "1", 
                                               "2+"),
                                    right = TRUE,
                                    ordered_result = TRUE))
Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-sibsp-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, SibSp_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

#### By Response

```{r eda-kruskal-sibsp-range-response}
(sjstats::means_by_group(Titanic.df, "Survived", "SibSp_Range"))
(kt <- kruskal.test(Survived ~ SibSp_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-sibsp-range-response}
TSRS <- xtabs(~ Survived + SibSp_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TSRS)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TSRS), 2))
```

```{r eda-mosaic-sibsp-range-response}
vcd::mosaic(TSRS, shade=TRUE, legend=TRUE)
vcd::assoc(TSRS, shade=TRUE)
```

#### By `Sex`

```{r eda-kruskal-sibsp-range-gender}
(sjstats::means_by_group(Titanic.df, "SibSp_Range", "Sex"))
(kt <- kruskal.test(Sex ~ SibSp_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-sibsp-range-gender}
TSRG <- xtabs(~ Sex + SibSp_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TSRG)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TSRG), 2))
```

```{r eda-mosaic-sibsp-range-gender}
vcd::mosaic(TSRG, shade=TRUE, legend=TRUE)
vcd::assoc(TSRG, shade=TRUE)
```

#### By `Age`

```{r eda-kruskal-sibsp-range-age}
(sjstats::means_by_group(Titanic.df, "Age", "SibSp_Range"))
(r <- cor.test(~ Age + as.numeric(SibSp_Range), data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", r$p.value < 0.05)))
```

```{r eda-boxplot-sibsp-range-age, fig.width=6}
Titanic.df %>%
  dplyr::select(Subset, Survived, SibSp_Range, Age) %>%
  ggplot2::ggplot(ggplot2::aes(x = SibSp_Range, y = Age, fill = SibSp_Range, group = SibSp_Range)) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(title = "Boxplot of Age by Sibling/Spouse Count", x = "Number of Siblings and/or Spouses", y = "Passenger Age")
```


### `Parch_Range` (extracted from `Parch`)

#### Summary

```{r feature-parch-range}
Titanic.df %<>%
    dplyr::mutate(Parch_Range = cut(Parch, 
                                    breaks = c(-Inf, 0, 1, Inf),
                                    labels = c("0", 
                                               "1", 
                                               "2+"),
                                    right = TRUE,
                                    ordered_result = TRUE))
Titanic.train <- Titanic.df %>%
  dplyr::filter(Subset == "Train") %>%
  dplyr::select(-Subset)

Titanic.test <- Titanic.df %>%
  dplyr::filter(Subset == "Test") %>%
  dplyr::select(-c(Survived, Subset))
```

```{r feature-parch-range-response}
Titanic.df %>%
    dplyr::group_by(Subset, Parch_Range) %>%
    dplyr::summarise(Count = dplyr::n(), .groups = "drop_last") %>%
    dplyr::mutate(Proportion = round(Count / sum(Count), 2))
```

#### By Response

```{r eda-kruskal-parch-range-response}
(sjstats::means_by_group(Titanic.df, "Survived", "Parch_Range"))
(kt <- kruskal.test(Survived ~ Parch_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-parch-range-response}
TPCS <- xtabs(~ Survived + Parch_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TPCS)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TPCS), 2))
```

```{r eda-mosaic-parch-range-response}
vcd::mosaic(TPCS, shade=TRUE, legend=TRUE)
vcd::assoc(TPCS, shade=TRUE)
```

#### By `Sex`

```{r eda-kruskal-parch-range-gender}
(sjstats::means_by_group(Titanic.df, "Parch_Range", "Sex"))
(kt <- kruskal.test(Sex ~ Parch_Range, data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", kt$p.value < 0.05)))
```

```{r eda-crosstab-parch-range-gender}
TPCG <- xtabs(~ Sex + Parch_Range, data = Titanic.train)
writeLines("Cross-tabluation frequencies: ")
ftable(TPCG)
writeLines("Cross-tabluation proportions: ")
ftable(round(prop.table(TPCG), 2))
```

```{r eda-mosaic-parch-range-gender}
vcd::mosaic(TPCG, shade=TRUE, legend=TRUE)
vcd::assoc(TPCG, shade=TRUE)
```

#### By `Age`

```{r eda-kruskal-parch-range-age}
(sjstats::means_by_group(Titanic.df, "Age", "Parch_Range"))
(r <- cor.test(~ Age + as.numeric(Parch_Range), data = Titanic.train))
writeLines(c("null hypothesis: there is no relationship between the variables",
             "alternative hypothesis: there is a relationship between the variables",
             sprintf("Significant (P-value < 0.05)? %s", r$p.value < 0.05)))
```

```{r eda-boxplot-parch-range-age, fig.width=6}
Titanic.df %>%
  dplyr::select(Subset, Survived, Parch_Range, Age) %>%
  ggplot2::ggplot(ggplot2::aes(x = Parch_Range, y = Age, fill = Parch_Range, group = Parch_Range)) +
  ggplot2::geom_boxplot() +
  ggplot2::labs(title = "Boxplot of Age by Parent/Child Count", x = "Number of Parents and/or Children", y = "Passenger Age")
```


### Latent Features in `Name`

```{r feature-name}
# Prepare data frame for extraction from the `Name` feature.
(Titanic.name <- dplyr::select(Titanic.df, PassengerId, Subset, Name))
```

```{r feature-name-explore}
# PRCE format Regex for formatting the names.
Titanic.name.format <- "^(.+?(?=,))(?>,\\s+)(.+?\\.)(?>\\s+)(.*?(?=\\t|$|\\s\\())(?|(?=\\t|$)()(?:.*?(?=$))|(?:\\s\\(?)(.*?(?=\\)))(?:.*?(?=$)))"
Titanic.name.matches <- regexec(
    pattern = Titanic.name.format,
    text = Titanic.name$Name,
    perl = TRUE # For PRCE conditionals.
)

# Generate the matches.
Titanic.name.components <- c("Source", "Surname", "Title", "Given Name (B)", "Given Name (A)")
(Titanic.name.features <- Titanic.name %>%
    dplyr::pull(Name) %>%
    regmatches(Titanic.name.matches) %>%
    as.data.frame(row.names = Titanic.name.components, col.names = Titanic.name$PassengerId) %>%
    tibble::rownames_to_column(var = "Component") %>%
    tidyr::pivot_longer(cols = -Component, names_to = "PassengerId", values_to = "Value") %>%
    dplyr::arrange(PassengerId, desc(Component)) %>%
    dplyr::mutate(PassengerId = as.integer(sub("X", "", PassengerId))))
```


### `Title` (extracted from `Name`)

```{r feature-name-title-extract}
# Prepare data frame for extraction from the `Name` feature.



```

### `Martial Status` (extracted from `Name`)


### `Forename` (extracted from `Name`)

### `Middlename` (extracted from `Name`)

<!-- Additional names besides first name -->

### `Surname` (extracted from `Name`)

