---
title: "Titanic (Kaggle Competition): Data Imputation Validation"
output: 
  # Output as a previewable HTML Notebook.
  html_notebook:
    theme: united
    highlight: tango
    df_print: paged
    toc: true
    toc_depth: 3
  # Output as an rmarkdown::pdf_document()
  pdf_document:
    keep_tex: true
    highlight: tango
    fig_caption: true
    df_print: kable
    toc: true
    toc_depth: 3
# pandoc -> pdfLaTeX rendering options:
fontsize: 11pt
geometry:
  - margin=1in
  - heightrounded
documentclass: scrartcl
papersize: a4
urlcolor: violet
toccolor: blue
---

```{r setup, include=FALSE}
# Set the root directory.
knitr::opts_knit$set(root.dir = here::here(""))

# Set the chunk options.
knitr::opts_chunk$set(
    ## ---- Formatting Options ----
    comment = "",
    include = TRUE,
    collapse = TRUE,
    echo = TRUE,
    strip.white = TRUE,
    
    ## ---- Output Options ----
    message = TRUE,
    warning = TRUE,
    error = TRUE,
    results = 'markup',
    
    ## ---- Display Options ----
    fig.height = 4,
    fig.width = 6
)

# Source loader file.
source(here::here("R/loader.R"))
load.data()
```

This report contains imputation details about the Titanic prediction models.

## Setup

Our investigation makes use of several R packages that made it easier to summarize, visualize, manipulate, model, and impute missing data. These packages are specified below.

```{r load-packages, include=TRUE, echo=FALSE}
library(magrittr) # Attached for the `%>%` (pipe) and `extract` functions.

loadNamespace("renv")    # Project dependency tracker for collaboration.
loadNamespace("here")    # Used for obtaining paths relative to project root.
loadNamespace("readr")   # Tidyverse data reader.
loadNamespace("tidyr")   # Tidyverse data wrangler.
loadNamespace("qqplotr") # For plotting quantile-quantile plots.
loadNamespace("ggmosaic") # For plotting mosaic plots.
loadNamespace("visdat")  # For visualization of data.
loadNamespace("ltm")     # For EDA.
loadNamespace("sjstats") # For EDA.
loadNamespace("vcd")     # For EDA.
loadNamespace("skimr")   # For finding statistical summaries.
loadNamespace("naniar")  # Handle NA values in a tabular format.
loadNamespace("simputation")  # Missing data imputation.
loadNamespace("e1071")   # For classification.
loadNamespace("randomForest") # For classification.
loadNamespace("neuralnet")    # For deep learning.
```

## Associations

```{r eda-assoc-spearman}
s <- cor(model.matrix(~ 0 + Age + Fare + Pclass + Embarked + Survived + SibSp + Parch, data = Titanic.train, method = "spearman", exact = FALSE))
s.p_mat <- ggcorrplot::cor_pmat(model.matrix(~ 0 + Age + Fare + Pclass + Embarked + Survived + SibSp + Parch, data = Titanic.train), method = "spearman", exact = FALSE)
ggcorrplot::ggcorrplot(corr = s, p.mat = s.p_mat, method = "circle", type = "full", title = "Spearman Correlation")
```

```{r eda-assoc-kendall}
s <- cor(model.matrix(~ 0 + Age + Fare + Pclass + Embarked + Survived + SibSp + Parch, data = Titanic.train, method = "pearson", exact = FALSE))
s.p_mat <- ggcorrplot::cor_pmat(model.matrix(~ 0 + Age + Fare + Pclass + Embarked + Survived + SibSp + Parch, data = Titanic.train), method = "pearson", exact = FALSE)
ggcorrplot::ggcorrplot(corr = s, p.mat = s.p_mat, method = "circle", type = "full", title = "Pearson Correlation")
```

## Data Imputation

### `Age`

#### Missingness Summary

```{r impute-summary-age}
Titanic.train %>% dplyr::select(Age) %>% naniar::miss_var_summary()
```

```{r impute-baseline-age}
# Get subset of missing Age records.
(Age.miss <- Titanic.train %>%
  dplyr::filter(is.na(Age)))

# Get subset of complete Age records
# and make train/validation subsets.
Age.complete <- Titanic.train %>% dplyr::filter(!is.na(Age))
n_samples <- floor(0.80 * nrow(Age.complete))
indices <- sample(1:nrow(Age.complete), size = n_samples)
(Age.complete %<>%
  dplyr::mutate(Subset = dplyr::if_else(dplyr::row_number() %in% indices, "Train", "Validation")) %>%
  dplyr::mutate(Response = ifelse(Subset == "Validation", NA, Age)))
```

#### Associated Features

```{r impute-corr-age}
writeLines("Correlation: ")
cor.test(~ Age + Fare, data = Titanic.df)
cor.test(~ Age + SibSp, data = Titanic.df)
cor.test(~ Age + Parch, data = Titanic.df)

writeLines("ANOVA: ")
sjstats::anova_stats(aov(Age ~ Name, data = Titanic.df))
sjstats::anova_stats(aov(Age ~ Pclass, data = Titanic.df))
sjstats::anova_stats(aov(Age ~ Embarked, data = Titanic.df))
sjstats::anova_stats(aov(Age ~ Ticket, data = Titanic.df))
sjstats::anova_stats(aov(Age ~ Survived, data = Titanic.df))
sjstats::anova_stats(aov(Age ~ factor(SibSp), data = Titanic.df))
sjstats::anova_stats(aov(Age ~ factor(Parch), data = Titanic.df))
```

#### Imputation Candidates

```{r impute-median-age}
impute.median.age <- function(x, data) {
    return(Age.complete %>%
        simputation::impute_median(x) %>%
        dplyr::mutate(Residuals = Age - Response) %>%
        dplyr::summarise(
            MSE = mean(Residuals**2, na.rm = TRUE),
            RMSE = sqrt(MSE),
            MAE = mean(abs(Residuals), na.rm = TRUE),
            n_miss = sum(is.na(Response))
        ) %>% dplyr::mutate(Formula = deparse(x)))
}

dplyr::bind_rows(list(
    impute.median.age(Response ~ Survived),
    impute.median.age(Response ~ Pclass),
    impute.median.age(Response ~ SibSp),
    impute.median.age(Response ~ Parch),
    impute.median.age(Response ~ Survived | Pclass)
)) %>% dplyr::arrange(MSE, RMSE, MAE)
```

```{r impute-lm-age}
impute.lm.age <- function(x, data) {
    return(Age.complete %>%
        simputation::impute_lm(x) %>%
        dplyr::mutate(Residuals = Age - Response) %>%
        dplyr::summarise(
            MSE = mean(Residuals**2),
            RMSE = sqrt(MSE),
            MAE = mean(abs(Residuals)),
            n_miss = sum(is.na(Response))
        ) %>% dplyr::mutate(Formula = deparse(x)))
}

dplyr::bind_rows(list(
    impute.lm.age(Response ~ Survived | Pclass),
    impute.lm.age(Response ~ Pclass),
    impute.lm.age(Response ~ Survived + Pclass + SibSp + Parch),
    impute.lm.age(Response ~ Pclass + SibSp + Parch + Sex | Survived),
    impute.lm.age(Response ~ Pclass + SibSp + Parch | Survived + Sex),
    impute.lm.age(Response ~ SibSp + Parch | Survived + Sex + Pclass),
    impute.lm.age(Response ~ SibSp + Parch + Fare | Survived + Sex + Pclass),
    impute.lm.age(Response ~ SibSp + Parch + Fare | Survived + Sex + Pclass)
)) %>% dplyr::arrange(MSE, RMSE, MAE)
```

```{r impute-rf-age}
impute.rf.age <- function(x, data) {
    return(Age.complete %>%
        simputation::impute_rf(x) %>%
        dplyr::mutate(Residuals = Age - Response) %>%
        dplyr::summarise(
            MSE = mean(Residuals**2),
            RMSE = sqrt(MSE),
            MAE = mean(abs(Residuals)),
            n_miss = sum(is.na(Response))
        ) %>% dplyr::mutate(Formula = deparse(x)))
}

dplyr::bind_rows(list(
    impute.rf.age(Response ~ Survived | Pclass),
    impute.rf.age(Response ~ Pclass),
    impute.rf.age(Response ~ Survived + Pclass + SibSp + Parch),
    impute.rf.age(Response ~ Pclass + SibSp + Parch + Sex | Survived),
    impute.rf.age(Response ~ Pclass + SibSp + Parch | Survived + Sex),
    impute.rf.age(Response ~ SibSp + Parch | Survived + Sex + Pclass),
    impute.rf.age(Response ~ SibSp + Parch + Fare | Survived + Sex + Pclass)
)) %>% dplyr::arrange(MSE, RMSE, MAE)
```


#### Imputation

### `Cabin`

#### Missingness Summary

```{r impute-summary-cabin}
Titanic.df %>% dplyr::select(Subset, Cabin) %>% dplyr::group_by(Subset) %>% naniar::miss_var_summary()
```

#### Association Features

```{r impute-glm-cabin}
Cabin.known <- Titanic.df %>% dplyr::mutate(Cabin = ifelse(is.na(Cabin), "Unknown", "Known"), Cabin = factor(Cabin))

writeLines("Logistic Regressions: ")
summary(glm(Survived ~ 0 + Cabin, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + Pclass, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + Sex, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + Age, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + SibSp, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + Parch, data = Cabin.known, family = "binomial"))
# summary(glm(Cabin ~ 0 + Ticket, data = Cabin.known, family = "binomial"))
summary(glm(Cabin ~ 0 + Embarked, data = Cabin.known, family = "binomial"))
```

#### Imputation Candidates

```{r impute-baseline-cabin}
# Get subset of missing records.
(Cabin.miss <- Titanic.df %>%
  dplyr::filter(is.na(Cabin)))

# Get subset of complete records
# and make train/validation subsets.
Cabin.complete <- Titanic.train %>% dplyr::filter(!is.na(Cabin))
n_samples <- floor(0.80 * nrow(Cabin.complete))
indices <- sample(1:nrow(Cabin.complete), size = n_samples)
(Cabin.complete %<>%
  dplyr::mutate(Subset = dplyr::if_else(dplyr::row_number() %in% indices, "Train", "Validation")) %>%
  dplyr::mutate(Response = ifelse(Subset == "Validation", NA, Cabin)))
```
```{r impute-other-cabin}
impute.unknown.cabin <- function(x) {
  result <- Cabin.complete %>%
    dplyr::mutate(Response = ifelse(Subset == "Validation", "Unknown", Response))
  result.table <- xtabs(~ Cabin + Response, result)
  result %>%
    dplyr::summarise(
        Accuracy = mean(Cabin == Response),
        Error = mean(Cabin != Response),
        n_miss = sum(is.na(Response))
    ) %>% dplyr::mutate(Formula = "Impute 'Unknown'") %>%
    return()
}

dplyr::bind_rows(list(
    impute.unknown.cabin()
)) %>% dplyr::arrange(Error, desc(Accuracy))
```

### `Fare`

#### Missingness Summary

```{r impute-summary-fare}
Titanic.df %>% dplyr::select(Subset, Fare) %>% dplyr::group_by(Subset) %>% naniar::miss_var_summary()
```

#### Associated Features

```{r impute-corr-fare}
writeLines("Correlation: ")
cor.test(~ Fare + Age, data = Titanic.df)
cor.test(~ Fare + SibSp, data = Titanic.df)
cor.test(~ Fare + Parch, data = Titanic.df)

writeLines("ANOVA: ")
sjstats::anova_stats(aov(Fare ~ Name, data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ Pclass, data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ Embarked, data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ Ticket, data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ Survived, data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ factor(SibSp), data = Titanic.df))
sjstats::anova_stats(aov(Fare ~ factor(Parch), data = Titanic.df))
```

#### Imputation Candidates

```{r impute-baseline-fare}
# Get subset of missing records.
(Fare.miss <- Titanic.df %>%
  dplyr::filter(is.na(Fare)))

# Get subset of complete records
# and make train/validation subsets.
Fare.complete <- Titanic.train %>% dplyr::filter(!is.na(Fare))
n_samples <- floor(0.80 * nrow(Fare.complete))
indices <- sample(1:nrow(Fare.complete), size = n_samples)
(Fare.complete %<>%
  dplyr::mutate(Subset = dplyr::if_else(dplyr::row_number() %in% indices, "Train", "Validation")) %>%
  dplyr::mutate(Response = ifelse(Subset == "Validation", NA, Fare)))
```


```{r impute-median-fare}
impute.median.fare <- function(x, data) {
    return(Fare.complete %>%
        simputation::impute_median(x) %>%
        dplyr::mutate(Residuals = Fare - Response) %>%
        dplyr::summarise(
            MSE = mean(Residuals**2, na.rm = TRUE),
            RMSE = sqrt(MSE),
            MAE = mean(abs(Residuals), na.rm = TRUE),
            n_miss = sum(is.na(Response))
        ) %>% dplyr::mutate(Formula = deparse(x)))
}

dplyr::bind_rows(list(
    impute.median.fare(Response ~ Survived),
    impute.median.fare(Response ~ Pclass),
    impute.median.fare(Response ~ SibSp),
    impute.median.fare(Response ~ Parch),
    impute.median.fare(Response ~ Survived | Pclass)
)) %>% dplyr::arrange(MSE, RMSE, MAE)
```

```{r impute-lm-fare}
impute.lm.fare <- function(x, data) {
    return(Fare.complete %>%
        simputation::impute_lm(x) %>%
        dplyr::mutate(Residuals = Fare - Response) %>%
        dplyr::summarise(
            MSE = mean(Residuals**2),
            RMSE = sqrt(MSE),
            MAE = mean(abs(Residuals)),
            n_miss = sum(is.na(Response))
        ) %>% dplyr::mutate(Formula = deparse(x)))
}

dplyr::bind_rows(list(
    impute.lm.fare(Response ~ Survived | Pclass),
    impute.lm.fare(Response ~ Pclass),
    impute.lm.fare(Response ~ Survived + Pclass + SibSp + Parch),
    impute.lm.fare(Response ~ Pclass + SibSp + Parch + Sex | Survived),
    impute.lm.fare(Response ~ Pclass + SibSp + Parch | Survived + Sex),
    impute.lm.fare(Response ~ SibSp + Parch | Survived + Sex + Pclass),
    impute.lm.fare(Response ~ SibSp + Parch + Fare | Survived + Sex + Pclass),
    impute.lm.fare(Response ~ SibSp + Parch + Fare | Survived + Sex + Pclass)
)) %>% dplyr::arrange(MSE, RMSE, MAE)
```


### `Embarked`

#### Missingness Summary

```{r impute-summary-embarked}
Titanic.df %>% dplyr::select(Subset, Embarked) %>% dplyr::group_by(Subset) %>% naniar::miss_var_summary()
```

#### Associated Features

```{r impute-corr-embarked}
writeLines("Kruskal-Wallis Tests: ")
kruskal.test(Age ~ Embarked, data = Titanic.df)
kruskal.test(SibSp ~ Embarked, data = Titanic.df)
kruskal.test(Parch ~ Embarked, data = Titanic.df)

writeLines("ANOVA: ")
summary(glm(Survived ~ 0 + Embarked, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ 0 + Pclass, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ 0 + Sex, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ 0 + Cabin, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ Age, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ SibSp, data = Cabin.known, family = "binomial"))
summary(glm(Embarked ~ Parch, data = Cabin.known, family = "binomial"))
```

#### Imputation Candidates

```{r impute-baseline-embarked}
# Get subset of missing records.
(Embarked.miss <- Titanic.df %>%
  dplyr::filter(is.na(Fare)))

# Get subset of complete records
# and make train/validation subsets.
Embarked.complete <- Titanic.train %>% dplyr::filter(!is.na(Embarked))
n_samples <- floor(0.80 * nrow(Embarked.complete))
indices <- sample(1:nrow(Embarked.complete), size = n_samples)
(Embarked.complete %<>%
  dplyr::mutate(Subset = dplyr::if_else(dplyr::row_number() %in% indices, "Train", "Validation")) %>%
  dplyr::mutate(Response = as.factor(Embarked), Response = factor(ifelse(Subset == "Validation", NA, Response), labels = c("S", "C", "Q"))))
```

```{r impute-constant-embarked}
impute.constant.embarked <- function(x) {
  result <- Embarked.complete %>%
    dplyr::mutate(Response = ifelse(Subset == "Validation", "S", Response))
  result.table <- xtabs(~ Embarked + Response, result)
  result %>%
    dplyr::summarise(
        Accuracy = mean(Embarked == Response),
        Error = mean(Embarked != Response),
        n_miss = sum(is.na(Response))
    ) %>% dplyr::mutate(Formula = "Impute 'Unknown'", Method = "Constant 'S'") %>%
    return()
}

dplyr::bind_rows(list(
    impute.constant.embarked()
)) %>% dplyr::arrange(Error, desc(Accuracy))
```

```{r impute-model-embarked}
impute.model.embarked <- function(x, data, model, model.name = deparse(substitute(model)), ...) {
  
  # Prepare model data (with certain fields temporarily imputed for testing purposes).
  data.train <- data %>%
    dplyr::filter(Subset == "Train") %>%
    dplyr::mutate(Age = dplyr::coalesce(Age, median(Age, na.rm = TRUE)), 
                  Fare = dplyr::coalesce(Fare, median(Fare, na.rm = TRUE)),
                  Cabin = dplyr::coalesce(Cabin, "Unknown"))
  
  data.test <- data %>%
    dplyr::filter(Subset == "Validation") %>%
    dplyr::mutate(Age = dplyr::coalesce(Age, median(Age, na.rm = TRUE)), 
                  Fare = dplyr::coalesce(Fare, median(Fare, na.rm = TRUE)), 
                  Cabin = dplyr::coalesce(Cabin, "Unknown")) %>% dplyr::select(-Response)
  
  # Fit imputation model.
  model.fit <- do.call(model, list(formula = x, data = data.train), quote = TRUE)
  
  # Make predictions.
  predictions <- dplyr::tibble(
    PassengerId = data.test %>% dplyr::pull(PassengerId),
    Response = factor(predict(model.fit, newdata = data.test))
  )

  # Impute missing values.
  result <- data %>%
    dplyr::rows_patch(y = predictions, by = "PassengerId")
  
  result.table <- xtabs(~ Embarked + Response, result)
  result %>%
    dplyr::summarise(
        Accuracy = mean(Embarked == Response),
        Error = mean(Embarked != Response),
        n_miss = sum(is.na(Response))
    ) %>% dplyr::mutate(Formula = deparse(x), Method = model.name) %>%
    return()
}

dplyr::bind_rows(list(
    impute.model.embarked(Embarked ~ Survived, Embarked.complete, randomForest::randomForest),
    impute.model.embarked(Embarked ~ Pclass + Sex + Survived, Embarked.complete, randomForest::randomForest),
    impute.model.embarked(Embarked ~ Pclass + Sex + Fare + SibSp + Parch, Embarked.complete, randomForest::randomForest),
    impute.model.embarked(Embarked ~ Pclass + Sex + Age + Fare + SibSp + Parch, Embarked.complete, randomForest::randomForest)
    
)) %>% dplyr::arrange(Error, desc(Accuracy))
```



